Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-11-22 paper/tex/discussion.tex
address more Seth typos

\section{Discussion and future work}
\label{sec:discussion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
In Section~\ref{sec:caution}, we reframe the strong performance of our pathological \pzpdf\ technique, \trainz, as a cautionary tale about the importance of choosing appropriate comparison metrics.
In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study
In Section~\ref{sec:futuredata}, we discuss the enhancements of the mock data set that will be necessary to enable the future experiments.

@@

We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments} and~\ref{sec:pointmetrics}.

@@

The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimats are constant as a function of true redshift, i.~e.~a horizontal line at the mode and mean of the training set distribution respectively.
The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.

% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.
%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
In summary, context is crucial to defend against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves our science goals}, not the one that performs best on a metric that does not reflect those goals.
In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.
% As stated earlier, science applications may be sensitive to individual \pzpdf s, \pzpdf\ ensembles, \pz\ point estimates, or science-specific statistics like $\hat{N}(z)$.
% We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
% The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
% It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
%\red{[the challenges faced in this project]}
%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}
%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}

@@

This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalogue-based photometry and representative training data, but the resilience of each code to such realistic imperfections in prior information has not yet been evaluated.
A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
Outright redshift failures due to emission line misidentification or noise spikes may be modeled by the inclusion of a small number of high-confidence yet false redshifts.
We plan to perform a full sensitivity analysis on a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which potential training set galaxies are most likely to be excluded.

Appendix~\ref{sec:moments} only addresses the stacked estimator of the redshift distribution of the entire galaxy catalogue rather than subsets in bins, tomographic or otherwise.
The effects of tomographic binning scheme will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.

@@

To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the \lsstdesc\ Science Roadmap, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
This initial paper explored a data set that was constructed at the catalogue level, with no inclusion of the complications that come from measuring photometry from images.
Future data challenges will move to catalogues constructed from mock images, including the complications of deblending, sensor inefficiencies, and heterogeneous observing conditions, all anticipated to affect the measured colours of \lsst's galaxy sample \citep{Dawson:2016}.

The DC1 galaxy SEDs were linear combinations of just five basis SED templates, but a next generation of data for \pzpdf\ investigations must include a broader range of physical properties.
Though we only considered $z < 2$ here, \lsst\ 10-year data will contain $z > 2$ galaxies, plagued by fainter apparent magnitudes and anomalous colours due to stellar evolution.
A subsequent study must also have a data set that includes low-level active galactic nuclei (AGN) features in the SEDs, which perturb colours and other host galaxy properties.
An observational degeneracy between the Lyman break of a $z \sim 2-3$ galaxy from the Balmer break of a $z \sim 0.2-0.3$ galaxy is a known source of catastrophic outliers \citep{Massarotti:2001} that was not effectively included in this study.
To gauge the sensitivity of \pzpdf\ estimators to catastrophic outliers, our data set must include realistic high-redshift galaxy populations.

>>>>

\section{Discussion and future work}
\label{sec:discussion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
In Section~\ref{sec:caution}, we reframe the strong performance of our pathological \pzpdf\ technique, \trainz, as a cautionary tale about the importance of choosing appropriate comparison metrics.
In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study.
In Section~\ref{sec:futuredata}, we discuss the enhancements of the mock data set that will be necessary to enable the future experiments.

@@

We remind the reader that codes utilized in this study were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments} and~\ref{sec:pointmetrics}.

@@

The CDE loss and point estimate metrics appropriately penalize \trainz's naivete.
As shown in Appendix~\ref{sec:pointmetrics}, \trainz ~has identical $ZPEAK$ and $ZWEIGHT$ values for every galaxy, and thus the \pz\ point estimates are constant as a function of true redshift, i.e.~a horizontal line at the mode and mean of the training set distribution respectively.
The explicit dependence on the individual posteriors in the calculation of the CDE loss, described in Section~\ref{sec:cdelossresults}, distinguishes this metric from those of the \pzpdf\ ensemble and stacked estimator of the redshift distribution, despite their prevalence in the \pz\ literature.

% While looking at one, or even most of our metrics would have given the impression that this estimator was nearly optimal, red flags in CDE loss and point estimates reveal the problems.
%[more caution on over-reliance on metrics, and making sure that metrics test all of what you want] \red{this definitely still needs work}
In summary, context is crucial to defend against deceptively strong performers such as \trainz; \textbf{the best \pzpdf\ method is the one that most effectively achieves the science goals of a particular study}, not the one that performs best on a metric that does not reflect those goals.
In the absence of a single scientific motivation or the information necessary for a principled metric definition, we must consider many metrics and be critical of the information transmitted by each.
% As stated earlier, science applications may be sensitive to individual \pzpdf s, \pzpdf\ ensembles, \pz\ point estimates, or science-specific statistics like $\hat{N}(z)$.
% We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.
% The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.
% It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
%\red{[the challenges faced in this project]}
%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}
%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}

@@

This initial paper explores \pzpdf\ code performance in idealized conditions with perfect catalogue-based photometry and representative training data, but the resilience of each code to realistic imperfections in prior information has not yet been evaluated.
A top priority for a follow-up study is to test realistic forms of incomplete, erroneous, and non-representative template libraries and training sets as well as the impact of other forms of external priors that must be ingested by the codes, major concerns in \citet{Newman:2015, Masters:2017}.
Outright redshift failures due to emission line misidentification or noise spikes may be modeled by the inclusion of a small number of high-confidence yet false redshifts.
We plan to perform a full sensitivity analysis on a realistically incomplete training set of spectroscopic galaxies, modeling the performance of spectrographs, emission-line properties, and expected signal-to-noise to determine which potential training set galaxies are most likely to be excluded.

Appendix~\ref{sec:moments} only addresses the stacked estimator of the redshift distribution of the entire galaxy catalogue rather than subsets in bins, tomographic or otherwise.
The effects of tomographic binning schemes will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.

@@

To make optimal use of the \lsst\ data for cosmological and other astrophysical analyses of the \lsstdesc\ Science Roadmap, future investigations that build upon this one will require a more sophisticated set of galaxy photometry and redshifts.
%\item \red{inclusion of photometric errors, image based effects, blending, lensing, sky, mask, etc...}
This initial paper explored a data set that was constructed at the catalogue level, with no inclusion of the complications that arise from photometric measurements of imaging data.
Future data challenges will move to catalogues constructed from mock images, including the complications of deblending, sensor inefficiencies, and heterogeneous observing conditions, all anticipated to affect the measured colours of \lsst's galaxy sample \citep{Dawson:2016}.

The DC1 galaxy SEDs were linear combinations of just five basis SED templates, and the next generation of data for \pzpdf\ investigations must include a broader range of physical properties.
Though we only considered $z < 2$ here, \lsst\ 10-year data will contain $z > 2$ galaxies, plagued by fainter apparent magnitudes and anomalous colours due to stellar evolution.
A subsequent study must also have a data set that includes low-level active galactic nuclei (AGN) features in the SEDs, which perturb colours and other host galaxy properties.
An observational degeneracy between the Lyman break of a $z \sim 2-3$ galaxy from the Balmer break of a $z \sim 0.2-0.3$ galaxy is a known source of catastrophic outliers \citep{Massarotti:2001} that was not effectively included in this study.
To gauge the sensitivity of \pzpdf\ estimators to catastrophic outliers, our data set must include realistic high-redshift galaxy populations.

### 2019-11-22 paper/tex/results.tex
address more Seth typos

\begin{figure*}
%\centering
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
\caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
The \pzpdf s of all codes share some features for the example galaxies due to physical colour degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
The diverse algorithms and implementations induce differences in small-scale structure and sensitivity to physical systematics.}
\label{fig:pz_examples}
\end{figure*}

The most striking differences between codes are the small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
The $\mathrm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
This observation is consistent with the findings of \citet[]{Malz:2018} that the piecewise constant parameterization underperforms in the presence of small-scale structures.

@@

The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for \metaphor\ indicate overly narrow \pzpdf s.
These five codes in particular have a free parameter for bandwidth, which may be responsible for this vulnerability, in spite of the opportunity for fine-tuning with perfect prior information.
\flexzboost's ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) played a key role in diagonalizing the QQ plot, indicating a common avenue for improvement in the approaches that share this type of parameter.
On the other hand, the three purely template-based codes, \bpz, \eazy, and \lephare, do not exhibit much systematic broadening or narrowing, which may indicate that complete template coverage effectively defends from these effects.

@@

Close inspection of the extremes at PIT values of 0 and 1 reveal spikes in the first and last bin of the PIT histogram for some codes in Figure~\ref{fig:pitqq}, corresponding to catastrophic outliers where the true redshift lies outside of the support of the $p(z)$.
The catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
As expected, \trainz\ achieves precisely the 0.0002 value expected of an ideal PIT distribution.
\annz, \flexzboost, \lephare, and \metaphor\ have notably high catastrophic outlier rates $> 0.02$, exceeding 100 times the ideal PIT rate, meriting further investigation.

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection in the differences in how each statistic weights the tails of the distribution.
Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
\label{fig:pit_stats}
\end{figure*}

Figure~\ref{fig:pit_stats} highlights the relative values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within a factor of two.

>>>>

\begin{figure*}
%\centering
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
\caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
All \pzpdf s have been scaled to the same peak value.
The \pzpdf s of all codes share some features for the example galaxies due to physical colour degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
The diverse algorithms and implementations induce differences in small-scale structure and sensitivity to physical systematics.}
\label{fig:pz_examples}
\end{figure*}

The most striking differences between codes are the small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins for $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
The $\mathrm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
This observation is consistent with the findings of \citet[]{Malz:2018} that the piecewise constant form underperforms other parameterizations in the presence of small-scale structures.

@@

The PIT histograms of \delight, \cmnn, \skynet, and \tpz\ feature an underrepresentation of extreme values, indicative of overly broad \pzpdf s, while the overrepresentation of extreme values for \metaphor\ indicates overly narrow \pzpdf s.
These five codes in particular have a free parameter for bandwidth, which may be responsible for this vulnerability, in spite of the opportunity for fine-tuning with perfect prior information.
\flexzboost's ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) played a key role in diagonalizing the QQ plot, indicating a common avenue for improvement in the approaches that share this type of parameter.
On the other hand, the three purely template-based codes, \bpz, \eazy, and \lephare, do not exhibit much systematic broadening or narrowing, which may indicate that complete template coverage effectively defends from these effects.

@@

Close inspection of the extremes at PIT values of 0 and 1 reveal spikes in the first and last bin of the PIT histogram for some codes in Figure~\ref{fig:pitqq}, corresponding to catastrophic outliers where the true redshift lies outside of the support of the $p(z)$.
The catastrophic outlier rates are provided in Table~\ref{tab:pitoutlier}.
As expected, \trainz\ achieves precisely the 0.0002 value expected of an ideal PIT distribution.
\annz, \flexzboost, \lephare, and \metaphor\ have notably high catastrophic outlier rates $> 0.02$, exceeding 100 times the ideal PIT rate, meriting further investigation elsewhere.

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection of the differences in how each statistic weights the tails of the distribution.
Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
\label{fig:pit_stats}
\end{figure*}

Figure~\ref{fig:pit_stats} highlights the relative values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with PIT values that more closely match the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within a factor of two.
